Limma Tutorial

### 2018-03-29 VirusHoneyBee/EdgeR/edgeR-btwnLane/makeDataMetrics.Rmd
Add all DEGs for VirusHoneyBee group

```{r}
minLib <- min(y$samples$lib.size)
keep <- rowSums(cpm(y)>4) >= 6
# Number of genes 15,314--> 8,672
y <- y[keep, , keep.lib.sizes=FALSE]
```

@@

```{r}
ggparcoord(data.frame(log(y[[1]]/colMeans(y[[1]]))), columns=1:24, alphaLines=0, boxplot=TRUE, scale="globalminmax") + coord_flip()
```

@@

```{r}
allGroups <- c(rep("N",12), rep("V",12))
y$samples$group <- allGroups

@@

```{r}
Group = factor(c(rep("N",12), rep("V",12)))
targets <- data.frame(rownames=paste0("Sample",1:24), Group = Group)
design <- model.matrix(~0+Group, data=y$samples)
colnames(design) <- levels(Group)
y <- estimateDisp(y, design)
plotBCV(y)
```

@@

    df <- y[[1]][rownames(y[[1]]) %in% rownames(lrt5), c(which(sapply(colnames(y[[1]]), function(x) strsplit(x, "[.]")[[1]][1]) %in% (colnames(fit)[i])), which(sapply(colnames(y[[1]]), function(x) strsplit(x, "[.]")[[1]][1]) %in% (colnames(fit)[j])))]
    firstLarger <- length(which(apply(df, 1, function(x) mean(x[1:6]) > mean(x[7:12]))))
    secondLarger <- length(which(apply(df, 1, function(x) mean(x[1:6]) < mean(x[7:12]))))

>>>>

```{r}
minLib <- min(y$samples$lib.size)
keep <- rowSums(cpm(y)>6) >= 3
# Number of genes 15,314--> 8,672
y <- y[keep, , keep.lib.sizes=FALSE]
```

@@

```{r}
ggparcoord(data.frame(log(y[[1]]/colMeans(y[[1]]))), columns=1:6, alphaLines=0, boxplot=TRUE, scale="globalminmax") + coord_flip()
```

@@

```{r}
allGroups <- c(rep("C",3), rep("T",3))
y$samples$group <- allGroups

@@

```{r}
Group = factor(c(rep("C",3), rep("T",3)))
targets <- data.frame(rownames=paste0("Sample",1:6), Group = Group)
design <- model.matrix(~0+Group, data=y$samples)
colnames(design) <- levels(Group)
y <- estimateDisp(y, design)
plotBCV(y)
```

@@

    df <- y[[1]][rownames(y[[1]]) %in% rownames(lrt5), c(which(sapply(colnames(y[[1]]), function(x) strsplit(x, "[.]")[[1]][1]) %in% (colnames(fit)[i])), which(sapply(colnames(y[[1]]), function(x) strsplit(x, "[.]")[[1]][1]) %in% (colnames(fit)[j])))]
    firstLarger <- length(which(apply(df, 1, function(x) mean(x[1:3]) > mean(x[4:6]))))
    secondLarger <- length(which(apply(df, 1, function(x) mean(x[1:3]) < mean(x[4:6]))))

### 2018-03-29 VirusHoneyBee/EdgeR/edgeR/makeDataMetrics.Rmd
Add all DEGs for VirusHoneyBee group

beeCounts <- readRDS("../../data/data.Rds")
beeCounts <- as.matrix(beeCounts)

@@

Next, I used edgeR normalization.
```{r}
y[[1]] <- betweenLaneNormalization(y[[1]], which="full", round=FALSE)
```

@@

```{r}
Group = factor(c(rep("C",3), rep("T",3)))
targets <- data.frame(rownames=paste0("Sample",1:6), Group = Group)
design <- model.matrix(~0+Group, data=y$samples)
colnames(design) <- levels(Group)
y <- estimateDisp(y, design)
plotBCV(y)
```

>>>>

beeCounts <- readRDS("../../data/data.Rds")

beeCounts <- as.matrix(beeCounts)

@@

Next, I used edgeR normalization.
```{r}
y <- calcNormFactors(y)
```

@@

```{r}
Group = factor(c(rep("C",3), rep("T",3)))
#targets <- data.frame(rownames=paste0("Sample",1:24), Group = Group)
design <- model.matrix(~0+Group, data=y$samples)
colnames(design) <- levels(Group)
y <- estimateDisp(y, design)
plotBCV(y)
```

### 2018-03-29 VirusHoneyBee/LimmaVoom/LimmaVoom-Virus/makeDataMetrics.Rmd
Add all DEGs for VirusHoneyBee group

thisPath <- getwd()
beeCounts <- readRDS("../data/data.Rds")
beeCounts <- as.matrix(beeCounts)
x <- DGEList(counts=beeCounts)

group <- factor(c(rep("N",12), rep("V",12)))
x$samples$group <- group
lane <- as.factor(c("L12","L12","L12","L12","L34","L34","L12","L12","L34","L34","L34","L34","L12","L12","L12","L12","L34","L34","L12","L12","L34","L34","L34","L34"))
x$samples$lane <- lane
```

@@

Make boxplots
```{r}
ggparcoord(data.frame(log(x[[1]]/colMeans(x[[1]]))), columns=1:24, alphaLines=0, boxplot=TRUE, scale="globalminmax") + coord_flip()
```

Create MDS plots
```{r}
library(RColorBrewer)
lcpm <- cpm(x, log=TRUE)
par(mfrow=c(1,2))
col.group <- group
levels(col.group) <-  brewer.pal(nlevels(col.group), "Set1")
col.group <- as.character(col.group)
col.lane <- lane
levels(col.lane) <-  brewer.pal(nlevels(col.lane), "Set2")
col.lane <- as.character(col.lane)
```

@@

```{r}
plotMDS(lcpm, labels=colnames(lcpm), col=col.lane, dim=c(3,4))
```

```{r}
glMDSPlot(lcpm, labels=paste(group, lane, colnames(lcpm), sep="_"), groups=x$samples[,c(1,4)], launch=FALSE)
```

Create design matrix. There are many ways to setup a design matrix. Here, we removed the intercept from group (the first factor), but kept the intercept from lane. This allows us to do contrasts with group more easily.
```{r}
design <- model.matrix(~0+group+lane)
colnames(design) <- gsub("group", "", colnames(design))

contr.matrix <- makeContrasts(
   N_V = N-V,
   levels = colnames(design))
```

>>>>

thisPath <- getwd()
beeCounts <- readRDS("../../data/data.Rds")
colnames(beeCounts) <- c(paste0(rep("C",3),".",1:3),paste0(rep("T",3),".",1:3))
beeCounts <- as.matrix(beeCounts)
x <- DGEList(counts=beeCounts)

group <- factor(c(rep("C",3), rep("T",3)))
x$samples$group <- group
```

@@

Make boxplots
```{r}
ggparcoord(data.frame(log(x[[1]]/colMeans(x[[1]]))), columns=1:6, alphaLines=0, boxplot=TRUE, scale="globalminmax") + coord_flip()
```

Create MDS plots
```{r}
library(RColorBrewer)
lcpm <- cpm(x, log=TRUE)
par(mfrow=c(1,2))
col.group <- group
levels(col.group) <-  brewer.pal(nlevels(col.group), "Set1")
col.group <- as.character(col.group)
```

@@

Create design matrix. There are many ways to setup a design matrix. Here, we removed the intercept from group (the first factor). This allows us to do contrasts with group more easily.
```{r}
design <- model.matrix(~0+group)
colnames(design) <- gsub("group", "", colnames(design))

contr.matrix <- makeContrasts(
   C_T = C-T,
   levels = colnames(design))
```

### 2018-03-29 VirusHoneyBee/LimmaVoom/makeDataMetrics.Rmd
Add all DEGs for VirusHoneyBee group

thisPath <- getwd()
beeCounts <- readRDS("../../data/data.Rds")
colnames(beeCounts) <- c(paste0(rep("C",3),".",1:3),paste0(rep("T",3),".",1:3))
beeCounts <- as.matrix(beeCounts)
x <- DGEList(counts=beeCounts)

>>>>

thisPath <- getwd()
beeCounts <- readRDS("../data/data.Rds")
beeCounts <- as.matrix(beeCounts)
x <- DGEList(counts=beeCounts)
